Best practices for forecasting changes in biodiversity: an example using breeding birds

### 2017-08-18 manuscript/draft.Rmd
30, not 32

In this paper, we attempt to forecast the species richness of breeding birds at
over 1,000 of sites located throughout North America, while following the best
practices in Box 1 for ecological forecasting. To do this, we combine 30 years
of time-series data on bird distributions from annual surveys with monthly
time-series of climate data and satellite-based remote-sensing. Datasets that
span a time scale of 30 years or more have only recently become available for
large-scale time-series based forecasting. A dataset of this size allows us to
model and assess changes a decade or more into the future in the presence of
shifts in environmental conditions on par with predicted climate change. We
compare traditional distribution modeling based approaches to spatial models of
species richness, time-series methods, and two simple baselines that predict
constant richness for each site, on average (Table 1). All of our forecasting
models account for uncertainty and observation error, are evaluated across
different time lags using hindcasting, and are publicly archived to allow future
assessment. We discuss the implications of these practices for our understanding
of, and confidence in, the resulting forecasts, and how we can continue to build
on these approaches to improve ecological forecasting in the future.

![Example predictions from six forecasting models for a single site. Data from
1982 through 2003, connected by solid lines, were used for training the models;
the remaining points were used for evaluating the models' forecasts. In each
panel, point estimates for each year are shown with lines; the darker ribbon
indicates the 68% prediction interval (1 standard deviation of uncertainty), and
the lighter ribbon indicates the 95% prediction interval. **A.** Single-site
models were trained independently on each site's observed richness values. The
first two models ("average" and "naive") served as baselines. **B.** The
environmental models were trained to predict richness based on elevation,
climate, and NDVI; the environmental models' predictions change from year to
year as environmental conditions change.](../figures/model_predictions.png)

>>>>

In this paper, we attempt to forecast the species richness of breeding birds at
over 1,000 of sites located throughout North America, while following the best
practices in Box 1 for ecological forecasting. To do this, we combine 32 years
of time-series data on bird distributions from annual surveys with monthly
time-series of climate data and satellite-based remote-sensing. Datasets that
span a time scale of 30 years or more have only recently become available for
large-scale time-series based forecasting. A dataset of this size allows us to
model and assess changes a decade or more into the future in the presence of
shifts in environmental conditions on par with predicted climate change. We
compare traditional distribution modeling based approaches to spatial models of
species richness, time-series methods, and two simple baselines that predict
constant richness for each site, on average (Table 1). All of our forecasting
models account for uncertainty and observation error, are evaluated across
different time lags using hindcasting, and are publicly archived to allow future
assessment. We discuss the implications of these practices for our understanding
of, and confidence in, the resulting forecasts, and how we can continue to build
on these approaches to improve ecological forecasting in the future.

![Example predictions from our six models for a single site. The data points
from 1982 through 2003, connected by solid lines, were used for training all six
models; the remaining points were used for evaluating the models' forecasts. In
each panel, point estimates for each year are shown with lines; the darker
ribbon indicates 1 standard deviation of uncertainty, and the lighter ribbon
indicates 95% confidence intervals. **A.** The single-site models models were
trained using one site at a time, based solely on observed richness values
shown. The first two models ("average" and "naive") served as our simple
forecasting baselines. **B.** The environmental models were trained to predict
richness based on elevation, climate, and NDVI; unlike the single-site models,
the environmental models' predictions change from year to year as environmental
conditions change.](../figures/model_predictions.png)
